---
title: "Exploratory Data Analysis"
author: "Chance Robinson"
date: "11/25/2019"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exploratory Data Analysis

## Library Imports


```{r library-imports, quietely=TRUE, warn.conflicts=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
# Correlation Matrix
library(ggcorrplot)
library(Hmisc)
```

## Load the CSV Data

```{r load-data}
data <- read.csv("../../data/CaseStudy2-data.csv", stringsAsFactors=TRUE, header = TRUE)
```

## Default Output

```{r head-data}
head(data)
```

### Identify Dimensions
 
```{r data-dimensions}
dim(data)
```


### Data Columns

```{r data-columns}
colnames(data)
```

### Convert Integers to Strings

* These numeric columns will be converted into strings 

```{r data-integers-to-strings}

data$ID <- as.character(data$ID)
data$EmployeeNumber <- as.character(data$EmployeeNumber)
data$EmployeeCount <- as.character(data$EmployeeCount)
data$StandardHours <- as.character(data$StandardHours)
data$Over18 <- as.character(data$Over18)

```


### Convert Integers to Factors

```{r data-integers-to-factors}

data$JobInvolvement <- factor(data$JobInvolvement, ordered = TRUE, 
                              levels = c(1, 2, 3, 4),
                              labels = c("Low", "Medium", "High", "Very High"))

data$JobSatisfaction <- factor(data$JobSatisfaction, ordered = TRUE, 
                              levels = c(1, 2, 3, 4),
                              labels = c("Low", "Medium", "High", "Very High"))

data$PerformanceRating <- factor(data$PerformanceRating, ordered = TRUE, 
                              levels = c(1, 2, 3, 4),
                              labels = c("Low", "Good", "Excellent", "Outstanding"))

data$RelationshipSatisfaction <- factor(data$RelationshipSatisfaction, ordered = TRUE, 
                              levels = c(1, 2, 3, 4),
                              labels = c("Low", "Medium", "High", "Very High"))

data$WorkLifeBalance <- factor(data$WorkLifeBalance, ordered = TRUE, 
                              levels = c(1, 2, 3, 4),
                              labels = c("Bad", "Better", "Good", "Best"))

### THIS WAS NOT ACTUALLY PROVIDED ON THE WALL
data$EnvironmentSatisfaction <- factor(data$EnvironmentSatisfaction, ordered = TRUE, 
                              levels = c(1, 2, 3, 4),
                              labels = c("Low", "Medium", "High", "Very High"))

data$StockOptionLevel <- factor(data$StockOptionLevel, ordered = TRUE, 
                              levels = c(0, 1, 2, 3),
                              labels = c("Zero", "One", "Two", "Three"))

data$JobLevel <- factor(data$JobLevel, ordered = TRUE, 
                              levels = c(1, 2, 3, 4, 5),
                              labels = c("One", "Two", "Three", "Four", "Five"))

data$Education <- factor(data$Education, ordered = FALSE, 
                              levels = c(1, 2, 3, 4, 5),
                              labels = c("One", "Two", "Three", "Four", "Five"))


```


### Describe the Data Types

```{r data-describe}
str(data)
```


### Numeric Columns

#### Column Names

```{r data-numeric}
data.numeric <- data %>%
  select_if(is.numeric)
```

#### Summary Tables

```{r data-numeric-summary}
summary(data.numeric)
```

```{r data-non-numeric-histograms}

# DistanceFromHome
# NumCompaniesWorked
# TotalWorkingYears
# TrainingTimesLastYear
# YearsAtCompany

ggplot(data=data.numeric, aes(data.numeric$YearsSinceLastPromotion)) + 
  geom_histogram()


```


### Non-Numeric Columns

#### Column Names

```{r data-non-numeric}
data.non.numeric <- data %>%
  select_if(is.factor)

colnames(data.non.numeric)
```


```{r data-non-numeric-frequency-table}

gather(data.non.numeric, "Column", "Value") %>%
  count(Column, Value) %>%
  group_by(Column) %>%             
  mutate(Percentage = prop.table(n)) %>%
  rename(Count = n)

```




```{r data-Attrition}
data %>%
  count(Attrition, sort = TRUE)

```


```{r data-business-travel}
data %>%
  count(BusinessTravel, sort = TRUE)
```

```{r data-department}
data.non.numeric %>%
  count(Department, sort = TRUE)
```


```{r data-education}
data.non.numeric %>%
  count(Education, sort = TRUE)
```


```{r data-education-field}
data.non.numeric %>%
  count(EducationField, sort = TRUE)
```

```{r data-gender}
data.non.numeric %>%
  count(Gender, sort = TRUE)
```

```{r data-job-role}
data.non.numeric %>%
  count(JobRole, sort = TRUE)
```

```{r data-marital-status}
data.non.numeric %>%
  count(MaritalStatus, sort = TRUE)
```


```{r data-over-time-1}
data.non.numeric %>%
  count(OverTime, sort = TRUE)
```

```{r data-over-time-2}
data %>%
  count(StandardHours, sort = TRUE)
```

### Correlation matrix for quantitative data

```{r corr-matrix}
# function for flattening and ordering the correlation matrix
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    correlation  =(cormat)[ut],
    p.value = pmat[ut]
    )
}
# See what variables are correlated with each other, p-values
correlation.matrix <- rcorr(as.matrix(data.numeric))
corDF <- data.frame(flattenCorrMatrix(correlation.matrix$r, correlation.matrix$P))
# Order the correlation matrix to show the highest correlated
# data.frame(corDF[order(-corDF$cor),])

quantDataModel <- corDF[which(abs(corDF$correlation) >= 0.5),]
data.frame(quantDataModel[order(-abs(quantDataModel$cor)),])

```

```{r corr.plot}

correlation.plot <- round(cor(data.numeric), 1)
ggcorrplot(correlation.plot, method = "square", type = "full", lab = TRUE)

```
