---
title: "Exploratory Data Analysis"
author: "Chance Robinson"
date: "11/25/2019"
output: 
  github_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Exploratory Data Analysis

## Library Imports


```{r library-imports, quietely=TRUE, warn.conflicts=FALSE}
library(tidyverse)
library(knitr)
library(kableExtra)
# Correlation Matrix
library(ggcorrplot)
library(Hmisc)
#Lasso
library(glmnet)
# Resampling
library(caret)
library(DMwR)
# KNN
library(class)
# Random Forrest
library(randomForest)  
library(e1071)  
```

## Load the CSV Data

```{r load-data}
data <- read.csv("../../data/CaseStudy2-data.csv", stringsAsFactors=TRUE)
```

## Default Output

```{r head-data}
head(data)
```

### Identify Dimensions
 
```{r data-dimensions}
dim(data)
```


### Data Columns

```{r data-columns}
colnames(data)
```

### Convert Integers to Strings

* These numeric columns will be converted into strings 

```{r data-integers-to-strings}

data$ID <- as.character(data$ID)
data$EmployeeNumber <- as.character(data$EmployeeNumber)
data$EmployeeCount <- as.character(data$EmployeeCount)
data$StandardHours <- as.character(data$StandardHours)
data$Over18 <- as.character(data$Over18)

```


### Convert Integers to Factors

```{r data-integers-to-factors}

data$JobInvolvement <- factor(data$JobInvolvement, ordered = TRUE, 
                              levels = c(1, 2, 3, 4),
                              labels = c("Low", "Medium", "High", "Very High"))

data$JobSatisfaction <- factor(data$JobSatisfaction, ordered = TRUE, 
                              levels = c(1, 2, 3, 4),
                              labels = c("Low", "Medium", "High", "Very High"))

data$PerformanceRating <- factor(data$PerformanceRating, ordered = TRUE, 
                              levels = c(1, 2, 3, 4),
                              labels = c("Low", "Good", "Excellent", "Outstanding"))

data$RelationshipSatisfaction <- factor(data$RelationshipSatisfaction, ordered = TRUE, 
                              levels = c(1, 2, 3, 4),
                              labels = c("Low", "Medium", "High", "Very High"))

data$WorkLifeBalance <- factor(data$WorkLifeBalance, ordered = TRUE, 
                              levels = c(1, 2, 3, 4),
                              labels = c("Bad", "Better", "Good", "Best"))

### THIS WAS NOT ACTUALLY PROVIDED ON THE WALL
data$EnvironmentSatisfaction <- factor(data$EnvironmentSatisfaction, ordered = TRUE, 
                              levels = c(1, 2, 3, 4),
                              labels = c("Low", "Medium", "High", "Very High"))

data$StockOptionLevel <- factor(data$StockOptionLevel, ordered = TRUE, 
                              levels = c(0, 1, 2, 3),
                              labels = c("Zero", "One", "Two", "Three"))

data$JobLevel <- factor(data$JobLevel, ordered = TRUE, 
                              levels = c(1, 2, 3, 4, 5),
                              labels = c("One", "Two", "Three", "Four", "Five"))

data$Education <- factor(data$Education, ordered = FALSE, 
                              levels = c(1, 2, 3, 4, 5),
                              labels = c("One", "Two", "Three", "Four", "Five"))


```



### Describe the Data Types

```{r data-describe}
str(data)
```


### Numeric Columns


#### Column Names

```{r data-numeric}
data.numeric <- data %>%
  select_if(is.numeric)
```

#### Summary Tables

```{r data-numeric-summary}
summary(data.numeric)
```


### Non-Numeric Columns

#### Column Names

```{r data-non-numeric}
data.non.numeric <- data %>%
  select_if(is.factor)

colnames(data.non.numeric)
```


```{r data-attrition}
data.non.numeric %>%
  count(Attrition, sort = TRUE)
```


```{r data-business-travel}
data %>%
  count(BusinessTravel, sort = TRUE)
```

```{r data-department}
data.non.numeric %>%
  count(Department, sort = TRUE)
```

```{r data-education-field}
data.non.numeric %>%
  count(EducationField, sort = TRUE)
```

```{r data-gender}
data.non.numeric %>%
  count(Gender, sort = TRUE)
```

```{r data-job-role}
data.non.numeric %>%
  count(JobRole, sort = TRUE)
```

```{r data-marital-status}
data.non.numeric %>%
  count(MaritalStatus, sort = TRUE)
```


```{r data-over-time}
data.non.numeric %>%
  count(OverTime, sort = TRUE)
```

```{r data-over-time}
data %>%
  count(StandardHours, sort = TRUE)
```

###Correlation matrix for quantitative data

```{r corr-matrix}
# function for flattening and ordering the correlation matrix
flattenCorrMatrix <- function(cormat, pmat) {
  ut <- upper.tri(cormat)
  data.frame(
    row = rownames(cormat)[row(cormat)[ut]],
    column = rownames(cormat)[col(cormat)[ut]],
    cor  =(cormat)[ut],
    p = pmat[ut]
    )
}
# See what variables are correlated with each other, p-values
correlation.matrix <- rcorr(as.matrix(data.numeric))
corDF <- data.frame(flattenCorrMatrix(correlation.matrix$r, correlation.matrix$P))
# Order the correlation matrix to show the highest correlated
# data.frame(corDF[order(-corDF$cor),])

quantDataModel <- corDF[which(corDF$cor >= 0.5),]
data.frame(quantDataModel[order(-quantDataModel$cor),])

```

```{r}
corr <- round(cor(data.numeric), 1)
  
ggcorrplot(corr, method = "square", type = "full", lab = TRUE)
```



```{r}

cols_to_remove <- c("ID", "EmployeeNumber", "EmployeeCount", "StandardHours", "Over18")

data.mod <- data %>%
  select(-cols_to_remove)


# data.mod.ds <- downSample(data.mod, data.mod$Attrition, list = FALSE)

data.mod.us <- upSample(data.mod, data.mod$Attrition, list = FALSE)

model.full <- glm(Attrition~., data = data.mod, family = binomial("logit"))
summary(model.full)


model.reduced.upsampled <- glm(Attrition~DistanceFromHome+JobInvolvement+OverTime+BusinessTravel+NumCompaniesWorked+WorkLifeBalance+EnvironmentSatisfaction+YearsSinceLastPromotion+YearsWithCurrManager+TrainingTimesLastYear, data = data.mod.us, family = binomial("logit"))
summary(model.reduced.upsampled)


```



```{r}
table(data.mod.us$Attrition)
```

```{r}

data.mod$predict <- as.factor(ifelse(model.full$fitted.values >0.5, "Yes", "No"))

data.mod.us$predict <- as.factor(ifelse(model.reduced.upsampled$fitted.values >0.5, "Yes", "No"))

mytable <- table(data.mod.us$Attrition, data.mod.us$predict)
rownames(mytable) <- c("Obs (No)", "Obs (Yes)")
colnames(mytable) <- c("Pred (No)","Pred (Yes)")
mytable


```


```{r}
# Confusion matrix
confusionMatrix(data.mod$Attrition, data.mod$predict, "Yes")
```

```{r}
# Confusion matrix
confusionMatrix(data.mod.us$Attrition, data.mod.us$predict, "Yes")
```


```{r}
set.seed(1234)
split.perc = .70


# DistanceFromHome+JobInvolvement+OverTime+BusinessTravel+NumCompaniesWorked+WorkLifeBalance+EnvironmentSatisfaction+YearsSinceLastPromotion+YearsWithCurrManager+TrainingTimesLastYear

# cols_to_keep = c("DistanceFromHome", "JobInvolvement", "OverTime", "BusinessTravel", "NumCompaniesWorked", "WorkLifeBalance", "EnvironmentSatisfaction", "YearsSinceLastPromotion", "YearsWithCurrManager", "TrainingTimesLastYear", "Attrition")

cols_to_keep = c("YearsSinceLastPromotion", "YearsWithCurrManager", "TrainingTimesLastYear", "DistanceFromHome", "NumCompaniesWorked", "YearsInCurrentRole", "TotalWorkingYears", "JobInvolvement", "OverTime", "BusinessTravel", "Attrition")



df_classify <- data.mod.us %>%
  select(cols_to_keep) %>%
  mutate(JobInvolvement = as.integer(JobInvolvement)) %>%
  mutate(OverTime = as.integer(OverTime)) %>%
  mutate(BusinessTravel = as.integer(BusinessTravel))


 #%>%
  # mutate(EnvironmentSatisfaction = as.integer(EnvironmentSatisfaction))

summary(df_classify)


# standardize values
df_classify$YearsSinceLastPromotion = scale(df_classify$YearsSinceLastPromotion)
df_classify$YearsWithCurrManager = scale(df_classify$YearsWithCurrManager)
df_classify$TrainingTimesLastYear = scale(df_classify$TrainingTimesLastYear)
df_classify$DistanceFromHome = scale(df_classify$DistanceFromHome)
df_classify$NumCompaniesWorked = scale(df_classify$NumCompaniesWorked)
df_classify$JobInvolvement = scale(df_classify$JobInvolvement)
df_classify$OverTime = scale(df_classify$OverTime)
df_classify$BusinessTravel = scale(df_classify$BusinessTravel)
df_classify$YearsInCurrentRole = scale(df_classify$YearsInCurrentRole)
df_classify$TotalWorkingYears = scale(df_classify$TotalWorkingYears)


train.indices = sample(1:dim(df_classify)[1],round(split.perc * dim(df_classify)[1]))

train = df_classify[train.indices,]
test = df_classify[-train.indices,]

dim(df_classify) # 939

dim(train) # 939
dim(test) # 939


```




```{r}

# original scale
classifications = knn(train[,c(1:10)],test[,c(1:10)], train$Attrition, prob = TRUE, k = 1)
confusionMatrix(table(test$Attrition,classifications), "Yes")

```

```{r}
rf = randomForest(Attrition~.,  
                   ntree = 100,
                   data = data.mod.us)
plot(rf) 


varImp(rf)

confusionMatrix(data=data.mod.us$predict,  
                reference=data.mod.us$Attrition, "Yes")

```




```{r}
iterations = 100
numks = 30


masterAcc = matrix(nrow = iterations, ncol = numks)

for(j in 1:iterations)
{
  train.indices = sample(1:dim(df_classify)[1],round(split.perc * dim(df_classify)[1]))
  train = df_classify[train.indices,]
  test = df_classify[-train.indices,]
  for(i in 1:numks)
  {
    classifications = knn(train[,c(1:10)],test[,c(1:10)], train$Attrition, prob = TRUE, k = i)
    table(test$Attrition,classifications)
    CM = confusionMatrix(table(test$Attrition,classifications), "Yes")
    masterAcc[j,i] = CM$overall[1]
  }
  
}

MeanAcc = colMeans(masterAcc)

plot(seq(1,numks,1),MeanAcc, type = "l")
```





### Linear Regression

```{r}

cols_to_remove <- c("ID", "EmployeeNumber", "EmployeeCount", "StandardHours", "Over18")

data.mod <- data %>%
  select(-cols_to_remove)

lm.model.full <- lm(MonthlyIncome~., data = data.mod)
summary(lm.model.full)

lm.model.reduced <- lm(MonthlyIncome~BusinessTravel+DailyRate+JobLevel+JobRole+TotalWorkingYears, data = data.mod)
summary(lm.model.reduced)


```



